An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-02-07 Articles/Article_EnerEcon_2014/EconGrowth4.Rnw
Edits to the text surrounding CES boundary mods.

The CES model without and with energy is given by 
Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$. 
The \texttt{cesEst} function of \texttt{micEconCES} provides 
several algorithm options for parameter estimation.
The default algorithm (Levenberg-Marquardt) does not respect
parameter constraints (see Section~\ref{sec:CES}) and, 
in our testing, nearly always violated them, often returning
negative values for elasticity of substitution parameters $\sigma$ and $\sigma_1$.
Thus, we used the two fitting algorithms available in \texttt{cesEst} that
respect coefficient constraints: \texttt{PORT} and \texttt{L-BFGS-B}.

Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho$ and $\rho_1$ (9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
which corresponds to $\sigma$ and $\sigma_1$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100, respectively.
During the grid search, values of $\rho$ and $\rho_1$ are fixed,
and values of $\lambda$, $\gamma$, $\delta$, and $\delta_1$ are estimated 
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms.
In all, 121 gradient searches in $\lambda$, $\gamma$, $\delta$, and $\delta_1$
at grid points representing all combinations of $\rho$ and $\rho_1$ are attempted.
During the grid search portion of our algorithm, starting values for the free parameters
are $\lambda$ = 0.015/year, $\delta$ = 0.5, $\delta_1$ = 0.5, 
and $\gamma$ is set to a value such that the mean of the residuals is zero
by the \texttt{cesEst} function.

Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B}) 
is attempted wherein all fitting parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are allowed to float. 
The start values for fitting parameters are taken from 
the grid search point that provided the lowest $sse$ (Equation~\ref{eq:sse}).

If resampled data are being fitted, a prior fit to historical data is available.
In the final step of our algorithm, 
a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
uses coefficients from the best 
fit to historical data as its starting point. 
In this final gradient search, all model parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are considered free parameters.

@@

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\delta$~=~0~or~1,
$\sigma_1$~=~0~or~$\infty$, and
$\sigma$~=~0~or~$\infty$.
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries~(rows~1--20).
Row~21 of Table~\ref{tab:CES_boundary_models} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from row~21 of Table~\ref{tab:CES_boundary_models}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are perfect complements,
and the Leontief model applies.%
  \footnote{
  E.g., when $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are perfect substitutes,
and the linear model applies.%
  \footnote{
  E.g., when $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

@@

The model with lowest $sse$ of all above trials 
is deemed the winning (i.e.,~best) model,
and its parameters are presented in graphs herein.
In special cases, two boundary models will be equivalent.
For example, boundary model~3 is identical to boundary model~1~(2)
if $x_1$ is always less than (greater than) $x_2$ for all years.
If two or more models have the same $sse$ (to six digits), 
the model with the lowest row number in Table~\ref{tab:CES_boundary_models} is deemed the winner,
thereby ensuring that the simplest model is selected.

>>>>

The CES model without and with energy is given by 
Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\gamma$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$. 
The \texttt{cesEst} function of \texttt{micEconCES} provides 
several algorithm options for parameter estimation.
The default algorithm (Levenberg-Marquardt) does not respect
parameter constraints (see Section~\ref{sec:CES}) and, 
in our testing, nearly always violated them, often returning
negative values for elasticity of substitution parameters $\sigma_1$ and $\sigma$.
Thus, we used the two fitting algorithms available in \texttt{cesEst} that
respect coefficient constraints: \texttt{PORT} and \texttt{L-BFGS-B}.

Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho_1$ and $\rho$ (9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
corresponding to $\sigma_1$ and $\sigma$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100, respectively.
During the grid search, values of $\rho_1$ and $\rho$ are fixed,
and values of $\gamma$, $\lambda$, $\delta_1$, and $\delta$ are estimated 
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms.
In all, 121 gradient searches in $\gamma$, $\lambda$, $\delta_1$, and $\delta$
at grid points representing all combinations of $\rho_1$ and $\rho$ are attempted.
During the grid search portion of our algorithm, starting values for the free parameters
are $\lambda$ = 0.015/year, $\delta_1$ = 0.5, $\delta$ = 0.5, 
and $\gamma$ is set by the \texttt{cesEst} function
to a value such that the mean of the residuals is zero.

Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B}) 
is attempted wherein all fitting parameters 
($\gamma$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
are allowed to float.
The starting values for fitting parameters are taken from 
the grid search point that provided the lowest $sse$ (Equation~\ref{eq:sse}).

If resampled data are being fitted, a prior fit to historical data is available.
In the final step of our algorithm, 
a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
uses coefficients from the best 
fit to historical data as its starting point. 
In this final gradient search, all model parameters 
($\gamma$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
are considered free parameters.

@@

In addition to the above trials, 
and in a manner similar to our approach 
to the Cobb-Douglas model~(see section \ref{sec:CD_Parameter_Estimation}),
we also fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\delta$~=~0~or~1,
$\sigma_1$~=~0~or~$\infty$, and
$\sigma$~=~0~or~$\infty$.
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries~(rows~1--20).
Row~21 of Table~\ref{tab:CES_boundary_models} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from row~21 of Table~\ref{tab:CES_boundary_models}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are perfect complements,
and the Leontief model applies.%
  \footnote{
  E.g., when $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$, row 3 in Table~\ref{tab:CES_boundary_models}.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are perfect substitutes,
and the linear model applies.%
  \footnote{
  E.g., when $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$,
  row 4 in Table~\ref{tab:CES_boundary_models}.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

@@

The model with lowest $sse$ of all above trials 
is deemed the winning (i.e.,~best) model,
and its parameters are presented in graphs herein.
In special cases, two boundary models will be equivalent.
For example, boundary model~3 is identical to boundary model~1~(2)
if $x_1$ is always less than (greater than) $x_2$ for all years.
If two or more models have the same $sse$ (to six digits), 
the model with the lowest row number in Table~\ref{tab:CES_boundary_models} 
is deemed the winning model,
thereby ensuring that the simpler model is selected.
